### 2020-07-17 paper.tex
Missing comma

\begin{algorithm*}[h]
\begin{algorithmic}
\Require probabilistic programs $\mathcal{P}$ and $\mathcal{Q}$; differentiable program $\mathcal{F}$; initial state $x$
\vspace{1mm}
\Procedure{auto-involutive-mcmc}{$\mathcal{P}$, $\mathcal{Q}$, $\mathcal{F}$, $x$}
    \State $(y, \log q_x(y)) \gets \textproc{trace-and-score}(\mathcal{Q}_x)$ \Comment{Sample $y \sim q_x(\cdot)$; compute log-density via probabilistic program}
    \State $(x' \oplus y', \log D) \gets \textproc{run-involution}(\mathcal{F}, x \oplus y)$ \Comment{{\small Compute $(x' \oplus y') = f(x \oplus y)$ and log(Radon-Nikodym derivative)}}
    \State $\log p(x) \gets \textproc{score}(\mathcal{P}, x)$ \Comment{Compute log-density via probabilistic program}
    \State $\log p(x') \gets \textproc{score}(\mathcal{P}, x')$ \Comment{Compute log-density via probabilistic program}
    \State $\log q_{x'}(y') \gets \textproc{score}(\mathcal{Q}_{x'}, y')$ \Comment{Compute log-density via probabilistic program}
    \State $\alpha \gets \min\{1, \exp(\log p(x') - \log p(x) + \log q_{x'}(y') - \log q_x(y) + \log D)\}$ \Comment{Compute acceptance probability}
    \State \textbf{with probability} \,$\alpha$ \,\Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure\\
\vspace{1mm}
\begin{minipage}[t]{0.49\linewidth}
\Procedure{trace-and-score}{$\mathcal{P}$}
   \State $\log w \gets 0$
   \State $x \gets \{\}$
   \execute[$\mathcal{P}$], but with $\mathtt{"}a \sim \mathrm{distribution}\mathtt{"} \equiv$ (
        \State 1. set $v \sim \mathrm{distribution}$
        \State 2. set $x[a] \gets v$
        \State 3. set $\log w \gets \log w + \textproc{logpdf}(\mathrm{distribution}, v)$
        \State 4. evaluate to $v$)
    \END
   \State \Return $(x, \log w)$
\EndProcedure
\end{minipage}%
\hfill
\begin{minipage}[t]{0.49\linewidth}
\Procedure{score}{$\mathcal{P}$, $x$}
   \State $\log w \gets 0$
   \execute[$\mathcal{P}$], but with $\mathtt{"}a \sim \mathrm{distribution}\mathtt{"} \equiv$ (
        \State 1. set $v \gets x[a]$
        \State 2. set $\log w \gets \log w + \textproc{logpdf}(\mathrm{distribution}, v)$
        \State 3. evaluate to $v$)
    \END
   \State \Return $\log w$
\EndProcedure
\end{minipage}\\
\vspace{1mm}
\Procedure{run-involution}{$\mathcal{F}$, $z$}
% TODO would be much simpler to describe if we didn't have to separate out x and y..
    \LineComment{Run the involution, and keep track of copied, read, and written continuous keys}
    \State $z' \gets \{\}$ \Comment{Initialize empty output dictionary}
    \execute[$\mathcal{F}$], but with
        \State $\mathtt{"write}(a, v)\mathtt{"} \equiv$
            (\algorithmicif\hspace{1mm} $a \in \mathcal{I}$
            \algorithmicthen\hspace{1mm} set $z'[a] \gets v$
            \algorithmicelse\hspace{1mm} set $z'[a] \gets v$ and set $W \gets W \cup \{a\})$
        \State $\mathtt{"read}(a)\mathtt{"} \equiv$
            (\algorithmicif\hspace{1mm} $a \in \mathcal{I}$
            \algorithmicthen\hspace{1mm} evaluate to $z[a]$
            \algorithmicelse\hspace{1mm} set $R \gets R \cup \{a\}$ and evaluate to $z[a]$)
        \State $\mathtt{"copy}(a_1, a_2)\mathtt{"} \equiv$ (set $z'[a_2] \gets z[a_1]$ and set $C \gets C \cup \{a_1\}$)
    \END
    \LineComment{Use automatic differentiation of $\mathcal{F}$ to compute Jacobian, skipping copied addresses}
    \For{$i$ in $1$ to $|W|$}
        \State $a_i \gets W_i$ \Comment{Pick $i$th key in $W$; the order does not matter}
        \State $J[i,:] \gets \nabla_{\mathbf{z}_{R \setminus C}} \left( h_{e(z)}[a_i] \right)$ \Comment{{\small Gradient of $z'[a_i] = h_{e(z)}[a_i]$ w.r.t. non-copied continuous inputs ($z_a$ for $a \in R \setminus C$)}}
    \EndFor
    \State \Return $(z', \log(|\det(J)|))$
\EndProcedure
\end{algorithmic}
\caption{Automated Involutive MCMC}
\label{alg:auto-involutive-mcmc}
\end{algorithm*}

@@

\paragraph{Automatic Sampling and Density Computation for Probabilistic Programs.}
Sampling traces from and computing densities of dictionaries under, the distribution on traces
induced by a probabilistic program is straight-forward to do, using a standard technique
in probabilistic programming.

>>>>

\begin{algorithm*}[h]
\begin{algorithmic}
\Require probabilistic programs $\mathcal{P}$ and $\mathcal{Q}$; differentiable program $\mathcal{F}$; initial state $x$
\vspace{1mm}
\Procedure{auto-involutive-mcmc}{$\mathcal{P}$, $\mathcal{Q}$, $\mathcal{F}$, $x$}
    \State $(y, \log q_x(y)) \gets \textproc{trace-and-score}(\mathcal{Q}_x)$ \Comment{Sample $y \sim q_x(\cdot)$; compute log-density via probabilistic program}
    \State $(x' \oplus y', \log D) \gets \textproc{run-involution}(\mathcal{F}, x \oplus y)$ \Comment{{\small Compute $(x' \oplus y') = f(x \oplus y)$ and log(Radon-Nikodym derivative)}}
    \State $\log p(x) \gets \textproc{score}(\mathcal{P}, x)$ \Comment{Compute log-density via probabilistic program}
    \State $\log p(x') \gets \textproc{score}(\mathcal{P}, x')$ \Comment{Compute log-density via probabilistic program}
    \State $\log q_{x'}(y') \gets \textproc{score}(\mathcal{Q}_{x'}, y')$ \Comment{Compute log-density via probabilistic program}
    \State $\alpha \gets \min\{1, \exp(\log p(x') - \log p(x) + \log q_{x'}(y') - \log q_x(y) + \log D)\}$ \Comment{Compute acceptance probability}
    \State \textbf{with probability} \,$\alpha$ \,\Return $x'$ \algorithmicelse \, \Return $x$  \algorithmicend
\EndProcedure\\
\vspace{1mm}
\begin{minipage}[t]{0.49\linewidth}
\Procedure{trace-and-score}{$\mathcal{P}$}
   \State $s \gets 0$
   \State $x \gets \{\}$
   \execute[$\mathcal{P}$], but with $\mathtt{"}a \sim \mathrm{distribution}\mathtt{"} \equiv$ (
        \State 1. set $v \sim \mathrm{distribution}$
        \State 2. set $x[a] \gets v$
        \State 3. set $s \gets s + \textproc{logpdf}(\mathrm{distribution}, v)$
        \State 4. evaluate to $v$)
    \END
   \State \Return $(x, s)$
\EndProcedure
\end{minipage}%
\hfill
\begin{minipage}[t]{0.49\linewidth}
\Procedure{score}{$\mathcal{P}$, $x$}
    \State $s \gets 0$
    \State $A \gets\{\}$
    \execute[$\mathcal{P}$], but with $\mathtt{"}a \sim \mathrm{distribution}\mathtt{"} \equiv$ (
        \State 1. set $v \gets x[a]$
        \State 2. set $A \gets A \cup \{a\}$
        \State 3. set $s \gets s + \textproc{logpdf}(\mathrm{distribution}, v)$
        \State 4. evaluate to $v$)
    \END
   \State \algorithmicif\hspace{1mm} $A \ne A_x$ \algorithmicthen\hspace{1mm} $s \gets -\infty$ \algorithmicend
   \State \Return $s$
\EndProcedure
\end{minipage}\\
\vspace{1mm}
\Procedure{run-involution}{$\mathcal{F}$, $z$}
% TODO would be much simpler to describe if we didn't have to separate out x and y..
    \LineComment{Run the involution, and keep track of copied, read, and written continuous keys}
    \State $z' \gets \{\}$ \Comment{Initialize empty output dictionary}
    \execute[$\mathcal{F}$], but with
        \State $\mathtt{"write}(a, v)\mathtt{"} \equiv$
            (\algorithmicif\hspace{1mm} $a \in \mathcal{I}$
            \algorithmicthen\hspace{1mm} set $z'[a] \gets v$
            \algorithmicelse\hspace{1mm} set $z'[a] \gets v$ and set $W \gets W \cup \{a\})$
        \State $\mathtt{"read}(a)\mathtt{"} \equiv$
            (\algorithmicif\hspace{1mm} $a \in \mathcal{I}$
            \algorithmicthen\hspace{1mm} evaluate to $z[a]$
            \algorithmicelse\hspace{1mm} set $R \gets R \cup \{a\}$ and evaluate to $z[a]$)
        \State $\mathtt{"copy}(a_1, a_2)\mathtt{"} \equiv$ (set $z'[a_2] \gets z[a_1]$ and set $C \gets C \cup \{a_1\}$)
    \END
    \LineComment{Use automatic differentiation of $\mathcal{F}$ to compute Jacobian, skipping copied addresses}
    \For{$i$ in $1$ to $|W|$}
        \State $a_i \gets W_i$ \Comment{Pick $i$th key in $W$; the order does not matter}
        \State $J[i,:] \gets \nabla_{\mathbf{z}_{R \setminus C}} \left( h_{e(z)}[a_i] \right)$ \Comment{{\small Gradient of $z'[a_i] = h_{e(z)}[a_i]$ w.r.t. non-copied continuous inputs ($z_a$ for $a \in R \setminus C$)}}
    \EndFor
    \State \Return $(z', \log(|\det(J)|))$
\EndProcedure
\end{algorithmic}
\caption{Automated Involutive MCMC}
\label{alg:auto-involutive-mcmc}
\end{algorithm*}

@@

\paragraph{Automatic Sampling and Density Computation for Probabilistic Programs.}
Sampling traces from, and computing densities of dictionaries under, the distribution on traces
induced by a probabilistic program is straight-forward to do, using a standard technique
in probabilistic programming.
